LaMachine: A meta-distribution for NLP software

### 2019-01-14 docs/lamachine_clarin2018.tex
paper update

LaMachine is a \emph{meta distribution} as it can be installed in various contexts. At its core, LaMachine consists of a
set of machine-parsable instructions on how to obtain, build (e.g. compile from source), install and configure software.
These are implemented using Ansible\footnote{\url{https://www.ansible.com}}.  This is notably different from the more
classical notion of Linux distributions, which generally provide their own repositories with (often binary) software
packages. LaMachine builds on this already established infrastructure by taking these repositories as a given and only
needs to know which repositories to use.  Similarly, there are different programming-language-specific ecosystems
providing their own repositories, such as the Python Package Index for Python, CRAN for R, CPAN for Perl, Maven Central
for Java.  LaMachine again relies on those to pull and install software from and never forks, archives, or modifies the
software in any way. In doing so, we compel participating software projects to adhere to well-established distribution
standards and ensure the software is more sustainable towards the future \cite{softwarequality}. Moreover, we ensure
that LaMachine never becomes a prerequisite for the software but merely a courtesy or convenience.

LaMachine provides ample flexibility that allows it to be deployable in different contexts. First of all there is
flexibility with regard to the target platform, where we support several major GNU/Linux distributions (Debian, Ubuntu,
CentOS, RedHat Enterprise Linux, Fedora, Arch Linux), as well as macOS (although with more limitations). Second, there is
flexibility with regard to the form, where we support \emph{containerisation} through
Docker\footnote{\url{https://www.docker.com}}, \emph{virtualisation} through Vagrant
and VirtualBox\footnote{\url{https://vagrant.org}, \url{https://www.virtualbox.org}}, direct remote provisioning through Ansible (for production
servers), or an installation that is either global to the machine or local in a custom directory for a specific user
(using \texttt{virtualenv}). Pre-built docker containers and virtual machine images with a limited selection of participating software are regularly
uploaded to the Docker Hub and Vagrant Cloud, respectively. The different flavours all offer a different degree of
seperation from the host OS, where Virtual Machines are completely virtualised, Docker Containers still share the kernel
with the host OS, and the machine-specific installation flavour actually compiles against the machine's distribution itself and thus
offers the least amount of overhead.

>>>>

LaMachine is a \emph{meta distribution} as it can be installed in various contexts. At its core, LaMachine consists of a
set of machine-parsable instructions on how to obtain, build (e.g. compile from source), install and configure software.
These are implemented using Ansible\footnote{\url{https://www.ansible.com}}.  This is notably different from the more
classical notion of Linux distributions, which generally provide their own repositories with (often binary) software
packages. LaMachine builds on this already established infrastructure by taking these repositories as a given and only
needs to know which repositories to use.  Similarly, there are different programming-language-specific ecosystems
providing their own repositories, such as the Python Package Index\footnote{\url{https://pypi.org}} for Python, CRAN\footnote{\url{https://cran.r-project.org/}} for R, CPAN\footnote{\url{https://www.cpan.org}} for Perl, Maven Central\footnote{\url{https://search.maven.org}} for Java.  As said, LaMachine relies on those to pull and install software from and never forks, archives, or modifies the software in any way. In doing so, we compel participating software projects to adhere to well-established distribution standards and ensure the software is more sustainable towards the future \cite{softwarequality}. Moreover, we ensure that LaMachine never becomes a prerequisite for the software but merely a courtesy or convenience.

LaMachine provides ample flexibility that allows it to be deployable in different contexts. First of all there is
flexibility regarding the installation form, which we call \emph{flavours}:

\begin{itemize}
    \item Local installation - This installs the bulk of LaMachine in a separate local virtual environment (a seperate
        directory) that has to be explicitly activated to be used. This also allows for multiple different installations
        of LaMachine on the same host system (e.g. for different users or with different software configurations).
        This local installation still actively relies on various global dependencies that are available through the
        package manager of your distribution.
    \item Global installation - This flavour is used for a host that is fully dedicated to LaMachine. Everything will be
        installed globally so there is only one installation possible, multiple users will all make use of the same
        installation.
    \item Docker container - Installs LaMachine in a container. Containerisation seperates the entire run-time
        environment
\end{itemize}

flexibility with regard to the target platform, where we support several major GNU/Linux distributions:

\begin{enumerate}
    \item Debian
    \item Ubuntu
    \item Arch Linux
    \item CentOS and RedHat Enterprise Linux
    \item Fedora
\end{enumerate}

In addition, we have limited support for macOS and the homebrew package manager\footnote{Alternatives such as macports are not
supported}.

(Debian, Ubuntu,
CentOS, RedHat Enterprise Linux, Fedora, Arch Linux), as well as macOS (although with more limitations). Second, there is
flexibility with regard to the form, where we support \emph{containerisation} through
Docker\footnote{\url{https://www.docker.com}}, \emph{virtualisation} through Vagrant
and VirtualBox\footnote{\url{https://vagrant.org}, \url{https://www.virtualbox.org}}, direct remote provisioning through Ansible (for production
servers), or an installation that is either global to the machine or local in a custom directory for a specific user
(using \texttt{virtualenv}). Pre-built docker containers and virtual machine images with a limited selection of participating software are regularly
uploaded to the Docker Hub and Vagrant Cloud, respectively. The different flavours all offer a different degree of
seperation from the host OS, where Virtual Machines are completely virtualised, Docker Containers still share the kernel
with the host OS, and the machine-specific installation flavour actually compiles against the machine's distribution itself and thus
offers the least amount of overhead.
